Determinism, Complexity, and Predictability in Computer Performance

### 2014-01-12 pecomplexity.tex
more text

@@

The relationship between prediction accuracy and the weighted permutation
entropy (WPE) is much as we conjectured: performance traces with high WPE ---
those whose temporal complexity is high, in the sense that the entropy
generation per value is approximately equal to the total entropy per value thus
little information is being propagated forward in time --- are indeed harder to
predict using the simple deterministic forecast model described in the previous
section.  The effects of changing the word length $n$ are also interesting:
using a longer word length generally lowers the WPE --- a natural consequence of
finite-length data --- but the falloff is less rapid in some traces than in
others, suggesting that those values are closer to the theoretical asymptote
that exists for perfect data. The persistent WPE values of 0.5--0.6 for the {\tt
col\_major} trace are consistent with dynamical chaos, further corroborating the
results of~\cite{mytkowicz09}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\textwidth]{figs/prediction_vs_entropy}
  \caption{The improvement of LMA over na\"{i}ve vs weighted permutation entropy.
  For each of these, the word length used is $5$.}
  \label{fig:pred_vs_ent}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\textwidth]{figs/SVD_wwpe}
  \caption{The weighted permutation entropy of one run of SVD. The gray bands
    are regions where the window overlaps regimes. The window size used is
    $5,000 \times 100,000$ instructions and the word length is $4$.}
  \label{fig:wwpe}
\end{figure}

>>>>

The weighted permutation entropy for the {\tt SVD} program is given in
Fig.~\cite{fig:wwpe}. To generate this image a window of 5,000 values slid over
the time series. Within each of those windows, the statistics over words of
length 4 are computed and the WPE is calculated. The gray bands denote regions
where the 5,000 value window overlapped visually-distinct regimes. It can be
seen that the behaviors of the weighted permutation entropy vary between
regimes.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1.0\textwidth]{figs/SVD_wwpe}
  \caption{The weighted permutation entropy of one run of SVD. The gray bands
    are regions where the window overlaps regimes. The window size used is
    $5,000 \times 100,000$ instructions and the word length is $4$.}
  \label{fig:wwpe}
\end{figure}

@@

The relationship between prediction accuracy and the weighted permutation
entropy (WPE) is much as we conjectured: performance traces with high WPE ---
those whose temporal complexity is high, in the sense that the entropy
generation per value is approximately equal to the total entropy per value thus
little information is being propagated forward in time --- are indeed harder to
predict using the simple deterministic forecast model described in the previous
section. See Fig.~\cite{fig:pred_vs_ent} for the distribution. The effects of
changing the word length $n$ are also interesting: using a longer word length
generally lowers the WPE --- a natural consequence of finite-length data --- but
the falloff is less rapid in some traces than in others, suggesting that those
values are closer to the theoretical asymptote that exists for perfect data. The
persistent WPE values of 0.5--0.6 for the {\tt col\_major} trace are consistent
with dynamical chaos, further corroborating the results of~\cite{mytkowicz09}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1.0\textwidth]{figs/prediction_vs_entropy}
  \caption{The improvement of LMA over na\"{i}ve vs weighted permutation entropy.
  For each of these, the word length used is $5$.}
  \label{fig:pred_vs_ent}
\end{figure}

@@
